Supplemental Materials

### 2024-07-08 products/manuscript/Manuscript_Final.qmd
Final update for draft
Draft for review

This study builds on an important body of work in economic history that uses hospital records to explore historical trends in birth weights [@costa1998; @costa2004; @goldin1989; @steckel1986; @steckel1998]. It also dialogues with studies in the human sciences that hypothesize an intergenerational effect of the poor health conditions, including poor nutrition and excessive workloads, experienced by enslaved individuals on their descendants across multiple generations [@jasienska2009; @leimert2020; @martins2024].

@@

[^1]: For complete information on the source, see @roth2020; @roth2020a. My Github repository, in particular, the `Supplemental_Materials` file, includes summary information on data processing, exploratory analysis, and model fitting. All materials, including the data set, are open access and reproducible, with instructions on the schematic of the workflow available online. See @roth2024.

Relying on the dataset, I ask whether the legacy of slavery, both as possible intergenerational inheritance patterns or as contemporary experiences of discrimination in a post-abolition society, affected the health of infants born in the first public maternity hospital in 1920s Brazil using birth weight as a proxy for general fetal and neonatal health. I quantify how maternal race, nationality, age, and parity explain racial disparities in infant birth weight.

@@

I use Ordinary Least Squares (OLS) to estimate the relationship of maternal variables on infant birth weight. In the supplemental analysis available online, I include results from logistic models that estimated the relationship of maternal variables on infant birth weight as a categorical variable (either normal birth weight, NBW, or low birth weight, LBW, according to current WHO classifications), and in line with some studies on birth weight [@worldhealthorganization2022; @falcao2020; @nyarko2013; @paixao2021; @silva1998; @wehby2015a; @wehby2017a]. Most logistic model results were insignificant, and thus I exclude them here. In my main models, I employ birth weight as a continuous variable in line with both current and historical studies [@andrasfay2020a; @costa1998; @costa2004; @david1997; @leal2006; @silveira2019]. Due to the lack of comprehensive information on gestational age in these data, I do not consider birth weight in reference to gestational age, as is best practice for disentangling pre-term birth (PTB) from low birth weight (LBW) [@cutland2017]. I discuss this constraint in the discussion below. The covariates include maternal age (recorded as a continuous variable in years), maternal skin color, gestational status, and maternal nationality.

>>>>

This study builds on an important body of work in economic history that uses hospital records to explore historical trends in birth weights [@costa1998; @costa2004; @goldin1989; @steckel1986; @steckel1998; @ward1993]. It also dialogues with studies in the human sciences that hypothesize an intergenerational effect of the poor health conditions, including poor nutrition and excessive workloads, experienced by enslaved individuals on their descendants across multiple generations [@jasienska2009; @leimert2020; @martins2024].

@@

[^1]: For complete information on the source, see @roth2020; @roth2020a. My Github repository, in particular, the `Supplemental_Materials` file, includes summary information on data processing, exploratory analysis, and model fitting. All materials, including the data set, are open access and reproducible, with instructions on the schematic of the workflow available online. See @roth2024. In the supplemental analysis, I further include results from logistic models that estimated the relationship of maternal variables on infant birth weight as a categorical variable (either normal birth weight, NBW, or low birth weight, LBW, according to current WHO classifications), and in line with some studies on birth weight @worldhealthorganization2022; @falcao2020; @nyarko2013; @paixao2021; @silva1998; @wehby2015a; @wehby2017a.

Relying on the dataset, I ask whether the legacy of slavery, both as possible intergenerational inheritance patterns or as contemporary experiences of discrimination in a post-abolition society, affected the health of infants born in the first public maternity hospital in 1920s Brazil using birth weight as a proxy for general fetal and neonatal health. I quantify how maternal skin color, nationality, age, and parity were associated with infant birth weight.

@@

I use Ordinary Least Squares (OLS) to estimate the relationship of maternal variables on infant birth weight. I employ birth weight as a continuous variable in line with both current and historical studies [@andrasfay2020a; @costa1998; @costa2004; @david1997; @leal2006; @silveira2019]. Due to the lack of comprehensive information on gestational age in these data, I do not consider birth weight in reference to gestational age, as is best practice for disentangling pre-term birth (PTB) from low birth weight (LBW) [@cutland2017]. I discuss this constraint in the discussion below. The covariates include maternal age (recorded as a continuous variable in years), maternal skin color, gestational status, and maternal nationality.

### 2024-07-08 products/manuscript/supplement/Supplemental_Materials.qmd
Final update for draft
Draft for review

---
title: "Supplemental Materials"
author: Cassia Roth
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
    bibliography: ../../../assets/Mada_Bib.bib
    csl: ../../../assets/chicago-fullnote-bibliography.csl
---
# Supplemental Materials

## Schematic of workflow

@@

In the exploratory analysis, I included recorded spontaneous abortions and all recorded live births. For the statistical models, I excluded both spontaneous abortions and any recorded live births weighing \<500 grams to avoid the inclusion of any possible stillbirths that were misrecorded as live births [@nyarko2013; @wehby2015; @wehby2017].

@@

```{r}
#| label: tbl-tablesupplemental
#| tbl-cap: "Supplemental Table"
#| echo: FALSE
table_supplemental <- readRDS(here::here("results", "tables", "table_supplemental_final.rds"))
table_gt <- gtsummary::as_gt(table_supplemental)
table_gt |> gt::as_raw_html()
```

@@

```{r}
#| label: tbl-table1
#| tbl-cap: "Model 1"
#| echo: FALSE
table1 <- readRDS(here::here("results", "tables", "table_fit1.rds"))
table_gt <- gtsummary::as_gt(table_fit1)
table_gt |> gt::as_raw_html()
```

@@

```{r}
#| label: tbl-table2
#| tbl-cap: "Model 2"
#| echo: FALSE
table2 <- readRDS(here::here("results", "tables", "table_fit2.rds"))
table_gt <- gtsummary::as_gt(table_fit2)
table_gt |> gt::as_raw_html()
```

@@

```{r}
#| label: tbl-table3
#| tbl-cap: "Model 3"
#| echo: FALSE
table3 <- readRDS(here::here("results", "tables", "table_fit3.rds"))
table_gt <- gtsummary::as_gt(table_fit3)
table_gt |> gt::as_raw_html()
```
In the first and third models, there appears to be an association between maternal skin color and infant birth weight, with Euro-descended women (the reference group) having infants with higher birth weights than Afro-descended women. In the second model, older mothers are associated with giving birth to infants with higher birth weights.

@@

```{r}
#| label: tbl-table4
#| tbl-cap: "Model 4"
#| echo: FALSE
table4 <- readRDS(here::here("results", "tables", "table_fit4.rds"))
table_gt <- gtsummary::as_gt(table_fit4)
table_gt |> gt::as_raw_html()
```

@@

```{r}
#| label: tbl-table5
#| tbl-cap: "Model 5"
#| echo: FALSE
table5 <- readRDS(here::here("results", "tables", "table_log1.rds"))
table_gt <- gtsummary::as_gt(table5)
table_gt |> gt::as_raw_html()
```

@@

```{r}
#| label: tbl-table6
#| tbl-cap: "Model 6"
#| echo: FALSE
table6 <- readRDS(here::here("results", "tables", "table_log2.rds"))
table_gt <- gtsummary::as_gt(table6)
table_gt |> gt::as_raw_html()
```

@@

```{r}
#| label: tbl-table7
#| tbl-cap: "Model 7"
#| echo: FALSE
table7 <- readRDS(here::here("results", "tables", "table_log3.rds"))
table_gt <- gtsummary::as_gt(table7)
table_gt |> gt::as_raw_html()
```

```{r}
#| label: tbl-table8
#| tbl-cap: "Model 8"
#| echo: FALSE
table8 <- readRDS(here::here("results", "tables", "table_log4.rds"))
table_gt <- gtsummary::as_gt(table8)
table_gt |> gt::as_raw_html()
```
In these two models, the only significant finding is the final logistic model, in which infants born to mixed-race mothers had 21% higher odds of being born with low birth weight compared to infants born to white mothers. The other models did not show significant results.

Best practice is to not categorize continuous variables, as this can include arbitrary cutoffs for variables that are best understood continuously [@Senn2005]. Thus, I will not include these logistic regressions in the final analysis.

@@

{{< pagebreak >}}

# References

>>>>

---
title: "Supplemental Materials"
author: Cassia Roth
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
    bibliography: ../../../assets/Mada_Bib.bib
    csl: ../../../assets/chicago-fullnote-bibliography.csl
    suppress-bibliography: true
---
# Supplemental Materials

```{r, echo=FALSE, message=FALSE}
#load packages
library(tidyverse) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(renv) #for package management
library(gtsummary) #for summary tables
library(flextable) #for nice tables
```
## Schematic of workflow

@@

In the exploratory analysis, I included recorded spontaneous abortions and all recorded live births. For the statistical models, I excluded both spontaneous abortions and any recorded live births weighing \<500 grams to avoid the inclusion of any possible stillbirths that were misrecorded as live births [@nyarko2013; @wehby2015a; @wehby2017a].

@@

```{r}
#| label: tbl-tablesupplemental
#| tbl-cap: "Supplemental Table"
#| echo: FALSE
table_supplemental <- readRDS(here::here("results", "tables", "table_supplemental_final.rds"))
flex_table <- as_flex_table(table_supplemental)
flex_table
```

@@

```{r}
#| label: tbl-table1
#| tbl-cap: "Model 1"
#| echo: FALSE
table1 <- readRDS(here::here("results", "tables", "table_fit1.rds"))
flex_table <- as_flex_table(table1)
flex_table
```

@@

```{r}
#| label: tbl-table2
#| tbl-cap: "Model 2"
#| echo: FALSE
table2 <- readRDS(here::here("results", "tables", "table_fit2.rds"))
flex_table <- as_flex_table(table2)
flex_table
```

@@

```{r}
#| label: tbl-table3
#| tbl-cap: "Model 3"
#| echo: FALSE
table3 <- readRDS(here::here("results", "tables", "table_fit3.rds"))
flex_table <- as_flex_table(table3)
flex_table
```
In the first and third models, there appears to be an association between maternal skin color and infant birth weight, with Euro-descended women (the reference group) having infants with higher birth weights than Afro-descended women. In the second model, older mothers are associated with giving birth to infants with higher birth weights.

@@

```{r}
#| label: tbl-table4
#| tbl-cap: "Model 4"
#| echo: FALSE
table4 <- readRDS(here::here("results", "tables", "table_fit4.rds"))
flex_table <- as_flex_table(table4)
flex_table
```

@@

```{r}
#| label: tbl-table5
#| tbl-cap: "Model 5"
#| echo: FALSE
table5 <- readRDS(here::here("results", "tables", "table_log1.rds"))
flex_table <- as_flex_table(table5)
flex_table
```

@@

```{r}
#| label: tbl-table6
#| tbl-cap: "Model 6"
#| echo: FALSE
table6 <- readRDS(here::here("results", "tables", "table_log2.rds"))
flex_table <- as_flex_table(table6)
flex_table
```

@@

```{r}
#| label: tbl-table7
#| tbl-cap: "Model 7"
#| echo: FALSE
table7 <- readRDS(here::here("results", "tables", "table_log3.rds"))
flex_table <- as_flex_table(table7)
flex_table
```

```{r}
#| label: tbl-table8
#| tbl-cap: "Model 8"
#| echo: FALSE
table8 <- readRDS(here::here("results", "tables", "table_log4.rds"))
flex_table <- as_flex_table(table8)
flex_table
```
In these two models, the only significant finding is the final logistic model, in which infants born to mixed-race mothers had 21% higher odds of being born with low birth weight compared to infants born to white mothers. The other models did not show significant results.

Best practice is to not categorize continuous variables, as this can include arbitrary cutoffs for variables that are best understood continuously [@senn2005]. Thus, I will not include these logistic regressions in the final analysis.

@@
